### 2018-07-02 README.Rmd
updated and read over README and vignette

The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat). The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.

There are two main commands in the `visdat` package:

@@

@@

There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity. These are described in detail in the vignette ["Experimental Features of visdat"](http://visdat.njtierney.com/articles/experimental_features.html).

@@

By default, `vis_dat` sorts the columns according to the type of the data in the vectors. You can turn this off by setting `sort_type = FALSE`.

```{r vis-dat-aq-sort-type}

vis_dat(airquality, 
        sort_type = FALSE)

### With many kinds of data

To demonstrate what visdat looks like when you have different kinds of data, we can look at the dataset `typical_data`, provided within `visdat`, and created with the excellent [`wakefield` package](https://github.com/trinker/wakefield).

```{r vis-dat-typical-data}

vis_dat(typical_data)

We can also look into using even wider data, looking at `typical_larger_data`

```{r vis-dat-typical-larger-data}

vis_dat(typical_data_large)

@@

We can explore the missing data further using `vis_miss()`.

@@

The percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.

You can cluster the missingness by setting `cluster = TRUE`.

@@

The columns can also just be arranged by columns with most missingness, by setting `sort_miss = TRUE`.

@@

`vis_miss` indicates when there is a very small amount of missing data at <0.1% missingness.

@@

`vis_miss` will also indicate when there is no missing data at all.

@@

@@

Thank you to Hadley Wickham for suggesting the use of the internals of `readr` to make `vis_guess` work.

Thank you to Miles McBain for his suggestions on how to improve `vis_guess`. This resulted in making it at least 2-3 times faster.

Thanks to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.

>>>>

`visdat` has 6 functions:

@@

- `vis_compare()` visualise differences between two dataframes of the same 
  dimensions

- `vis_expect()` visualise where certain conditions hold true in your data

- `vis_cor()` visualise the correlation of variables in a nice heatmap

- `vis_guess()` visualise the individual class of earch value in your data

@@

@@

@@

We can explore the missing data further using `vis_miss()`:

@@

Percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.

You can cluster the missingness by setting `cluster = TRUE`:

@@

Columns can also be arranged by columns with most missingness, by setting
`sort_miss = TRUE`:

@@

`vis_miss` indicates when there is a very small amount of missing data at <0.1%
missingness:

@@

`vis_miss` will also indicate when there is no missing data at all:

@@

To further explore the missingness structure in a dataset, I recommend the
[`naniar`](https://github.com/njtierney/naniar) package, which provides more
general tools for graphical and numerical exploration of missing values.

## Using `vis_compare()`

Sometimes you want to see what has changed in your data. `vis_compare()` displays the differences in two dataframes of the same size. Let's look at an example.

Let's make some changes to the `chickwts`, and compare this new dataset:

```{r vis-compare-iris}
chickwts_diff <- chickwts
chickwts_diff[sample(1:nrow(chickwts), 30),sample(1:ncol(chickwts), 2)] <- NA

vis_compare(chickwts_diff, chickwts)

Here the differences are marked in blue.

If you try and compare differences when the dimensions are different, you get 
an ugly error:

```{r vis-compare-error, eval = FALSE}

chickwts_diff_2 <- chickwts
chickwts_diff_2$new_col <- chickwts_diff_2$weight*2

vis_compare(chickwts, chickwts_diff_2)
# Error in vis_compare(chickwts, chickwts_diff_2) : 
#   Dimensions of df1 and df2 are not the same. vis_compare requires dataframes of identical dimensions.
```

## Using `vis_expect()`

`vis_expect` visualises certain conditions or values in your data. For example,
If you are not sure whether to expect values greater than 25 in your data
(airquality), you could write: `vis_expect(airquality, ~.x >= 25)`, and you can
see if there are times where the  values in your data are greater than or equal
to 25:

```{r vis-expect}

vis_expect(airquality, ~.x >= 25)

This shows the proportion of times that there are values greater than 25, as
well as the missings.

## Using `vis_cor()`

To make it easy to plot correlations of your data, use `vis_cor`:

```{r vis-cor}

vis_cor(airquality)

## Using `vis_guess()`

`vis_guess()` takes a guess at what each cell is. It's best illustrated using
some messy data, which we'll make here:

```{r create-messy-vec}

messy_vector <- c(TRUE,
                  T,
                  "TRUE",
                  "T",
                  "01/01/01",
                  "01/01/2001",
                  NA,
                  NaN,
                  "NA",
                  "Na",
                  "na",
                  "10",
                  10,
                  "10.1",
                  10.1,
                  "abc",
                  "$%TG")

set.seed(1114)
messy_df <- data.frame(var1 = messy_vector,
                       var2 = sample(messy_vector),
                       var3 = sample(messy_vector))

```{r vis-guess-messy-df, fig.show='hold', out.width='50%'}

vis_guess(messy_df)
vis_dat(messy_df)

So here we see that there are many different kinds of data in your dataframe. As
an analyst this might be a depressing finding. We can see this comparison above.

@@

Thank you to Hadley Wickham for suggesting the use of the internals of `readr`
to make `vis_guess` work. Thank you to Miles McBain for his suggestions on how
to improve `vis_guess`. This resulted in making it at least 2-3 times faster.
Thanks to Carson Sievert for writing the code that combined `plotly` with
`visdat`, and for Noam Ross for suggesting this in the first place. Thank you
also to Earo Wang and Stuart Lee for their help in getting capturing expressions
in `vis_expect`.
